---
title: "shinyBTC"
author: "Jan Gorecki"
date: '2014-11-11'
output:
  html_document: default
---

[shinyBTC](https://github.com/jangorecki/shinyBTC) app acts like a simple GUI for [Rbitcoin](https://github.com/jangorecki/Rbitcoin) package. Not all of Rbitcoin features were used in the app.

Use it wise, be aware of [MIT](http://opensource.org/licenses/MIT) license on which this app is released.

---

Rbitcoin introduction vignette
```{r, eval=FALSE}
vignette("introduction", package="Rbitcoin")
```

Any currency pair and methods are supported via `market.api.query` but with no translation to common interface.  
Below is a list of market + currency pairs supported via `market.api.process` which includes translation to common interface (pre/post processing API calls). API dictionary can be extended by user via `options`.
```{r}
getOption("Rbitcoin.api.dict")[!is.na(base), .(market, currency_pair = paste0(base,quote))][,unique(.SD)]
```

Details of host
```{r}
data.table(attribute = names(Sys.info()), value = Sys.info())
```

App instance start time
```{r}
Sys.time()
```

---
